---
title: "CDMOLBIO_gr20"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CDMOLBIO_gr20}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
This package replicates the central dogma of molecular biology, translating DNA into an amino acid sequence and visualizes the distribtuion of amino acid residues in the protein resulting from an AA sequence.

From GitHub repo `https://github.com/rforbiodatascience25/group_20_package`

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Examples
```{r setup}
library(CDMOLBIO)
```
### Generate random sequence with `generate_dna`
```{r}
generate_dna(10)
```

### Translate with `transcribe_dna_to_arn`
```{r}
transcribe_dna_to_arn("ATGCTGTGC")
```
### Extract codons to list from string with `get_codons`
```{r}
get_codons("AUGCUGUGCAUGCUGAUG")
```
### Take list of codons and return string of amino acids with `translation`
```{r}
translation(c("AUG", "UUU", "UUC"))
```
### Plot amino acid distribution with `aa_bar_plot`
```{r}
aa_bar_plot("MTGCTGAUMN")
```

### Using everything together
```{r}
generate_dna(12) |> transcribe_dna_to_arn() |> get_codons(start = 2) |> translation() |> aa_bar_plot()
```

## Use case
This package can be used to generate a random DNA sequence and follow the cetral dogma om molecular biology to see the resulting protein sequence. Furthermore it can visualize the composition of amino acids in the resulting protein. Functionality taking reading frame into account can be included and enable loading user defined sequences.

